---
title: "Working with Survey Monkeys"
author: "Rika Gorn"
date: "February 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(socviz)
library(glue)
library(knitr)
library(scales)
library(kableExtra)
library(dplyr)
library(tidyr)
library(readxl)
library(janitor)
library(ggplot2)
library(ggthemes)
library(forcats)
library(corrplot)
library(stringr)
library(tidyverse)
```

## R Markdown

As a data practicioner in the public sector, I frequently work on data compiled through surveys. We use surveys to get a feel for how clients are reacting to various programming, to analyze qualitative and quantitative data, and even to compile demographics. 90% of the time, the data is collected through [SurveyMonkey] (www.surveymonkey.com) - an online tool that works pretty well, especially if you upgrade to a paid version. As with any online software tool that optimizes UI through point and click tools, it has a few quirks. If, like me, you want to do the analysis on your own or in some other tool like R,I've found that there are a handful of best practices that can help.  I wanted to write a blogpost that went through some ways to work through these quirks, summarize likert style data, and create some easy, quick, and pretty visualizations.

### Exporting SM data

In the **Analyze Results** tab, SM allows users to download their data in a variety of ways. You can download the following in csv, xlsx, pdf, or spss outputs: 

*All summary data - I have not found this useful for my own analysis. This simply spits out the aggregate results of each question.

*All responses data - exports the raw dataset with individual responses as rows and questions as columns. This is what I primarily use for my analysis.

*All individual responses - exports all the answers for a specific individual.

### Cleaning data

In a standard survey, I will usually have a mix of demographic, binary (Yes or No), and likert scale questions. When I load my data in R, however, I can quickly see that SM has a funny way of organizing the questions and their associated labels.



```{r warning = FALSE}
mysurvey <- read_csv("mysurvey.csv")
glimpse(mysurvey)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
